### 2022-11-05 talk/talk.qmd
tweak thresholds and counting in topics over time

---
format: 
    revealjs:
        scrollable: true
        width: 1.1*1440
        height: 1080
        toc: false
        toc-depth: 1
        toc-title: Overview
        slide-number: true
        theme: [default, custom.scss]

title: "Mainstreaming Scientific Racism"
author: "Dan Hicks & Emilio Lobato"
institute: "University of California, Merced<br><https://mainstream-race-science.netlify.app>"
bibliography: ../paper/race_science.yaml
---

# Introduction {visibility="uncounted"}

## The return(?) of race science

:::: columns
::: column
![](img/Saini.png){width="70%"}\
@SainiSuperiorReturnRace2019
:::
::: column
> [Many people assumed that the eugenicists had all but disappeared with the Nazi regime, and that race science was pretty much finished at the same time.]{.blue .fragment}

> [Intellectual racism has always existed, and indeed for a chunk of history, it thrived. I believe it is still the toxic little seed at the heart of academia. However dead you might think it is, it needs only a little water, and now it’s raining.]{.red .fragment} 
:::
::::

## Institutions of race science

Pioneer Fund
: Founded 1937, during the decline of (institutionalized) eugenics\
  [@MehlerFoundationFascismNew1989; @TuckerFundingScientificRacism2002]

*Mankind Quarterly*
: Founded 1960 by marginalized academics as part of the backlash to *Brown v Board* [@MehlerFoundationFascismNew1989; @SchafferScientificRacismAgain2007; @JacksonScienceSegregationRace2005 ch. 7]

. . .

\

The Tobacco Strategy
: Used by regulated industries to create the appearance of legitimate scientific controversy and thereby block or delay regulation\
  [@OreskesMerchantsDoubtHow2011; @FernandezPintoKnowBetterNot2017]

## Echo chambers

- "creation and distribution of non-peer reviewed journals and pamphlets"
- "funding and organizing of symposiums" for industry-funded research
- ["In this way, the tobacco industry circumvented peer review standards in publication"]{.red} [@FernandezPintoKnowBetterNot2017]

::: {.incremental}
- "a bounded, enclosed media space that has the potential to both magnify the messages delivered within it and [insulate them from rebuttal]{.red}" [@JamiesonEchoChamberRush2008 76]
:::

[**Did Pioneer and *Mankind Quarterly* nurture race science by protecting it from mainstream peer review?**]{.fragment}\
[**Our analysis indicates *MQ* did not play this role**]{.fragment .red}

<!-- 
## PF and MQ as a race science echo chamber?

- Empirical, text-mining approach
- Identify Pioneer-funded researchers
- Identify journals that published their work
- Analyze all articles published in those journals, 1960-2010

\
 -->

<!-- 
# Race science, 1910-1960

```{r setup}
library(readr)
library(here)
library(plotly)
library(gt)

vis_dir = here('out')
```

<!--
## Some definitions

scientific racism
 ~ (purportedly) justifying racial inequality and colonialism by appealing to the epistemic authority of science

race science - wide sense
 ~ scientific research that is amenable to scientific racism

race science - narrow sense
 ~ scientific research that itself engages in scientific racism

- Ex: genotype clustering studies [@RosenbergGeneticStructureHuman2002] or polygenic scores for educational attainment [@HardenGeneticAssociationsMathematics2019] 
can be wide-sense race science without being narrow-sense race science [@CarlsonQuantifyingContextualizingImpact2020; WillsAreClustersRaces2017]

- This talk will focus on narrow-sense race science
-->

<!--
## Rise and fall of eugenics {.smaller}

:::: columns 
::: {.column width="33%"}
![[Charles Davenport (1866-1944)](https://en.wikipedia.org/wiki/Charles_Davenport), circa 1929. [Source](https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg)](img/Charles_Benedict_Davenport.jpg){width="80%"}
:::

::: {.column width="33%"}
![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. [Source](https://en.wikipedia.org/wiki/File:FranzBoas.jpg)](img/FranzBoas.jpg){width="80%"}
:::

::: {.column width="33%"}
![[Margaret Mead (1901-1978)](https://en.wikipedia.org/wiki/Margaret_Mead), 1948. [Source](https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg)](img/Margaret_Mead_(1901-1978).jpg){width="80%"}
:::
::::

- 1910: Davenport founds Eugenics Records Office at Cold Spring Harbor
- 1926: NRC "Committee on the Negro" includes both Davenport and Boas: "neither was influential enough to veto his rival's participation" [@BarkanRetreatScientificRacism1992 113]
- 1930s: *Coming of Age in Samoa* [@MeadComingAgeSamoa1928] popularizes cultural anthropology, "free[ing] anthropology from the shackles of biology" [@BarkanRetreatScientificRacism1992 134]
- 1940: Carnegie Institute defunds Eugenics Records Office
<!-- - 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth" [@BeagleholeRaceQuestion1950; though see @BrattainRaceRacismAntiracism2007] -->

<!--
## The decline of eugenics and the Pioneer Fund {.smaller}

:::: columns
::: {.column width="30%"}
![[Wickliffe Draper (1891-1972)](https://en.wikipedia.org/wiki/Wickliffe_Preston_Draper), textile heir, eugenicist, segregationist, 1918. [Source](https://en.wikipedia.org/wiki/File:Wickliffe_Draper.jpg)](img/Draper.jpg)
:::

::: {.column width="30%"}
![[Harry Laughlin (1880-1943)](https://en.wikipedia.org/wiki/Harry_Laughlin), 
superintendent of Eugenics Records Office, anti-immigrationist, c. 1929. [Source](https://en.wikipedia.org/wiki/File:Harry_Laughlin.jpg)](img/Harry_Laughlin.jpg){width="100%"}
:::

::: {.column width="30%"}
![[Frederick Osborn (1889-1981)](https://en.wikipedia.org/wiki/Frederick_Osborn), railroad heir, eugenicist, shaped postwar pivot to "positive eugenics," undated. [Source](https://www.meehanreports.com/eugen-PT-pt1.htm)](img/osbornphoto.jpg){width="100%"}
:::
::::

Founded in 1937 to "support academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 21, quoting Laughlin]
<!-- 
- Presidents
	- Harry Laughlin, 1937-1941
	- Frederick Osborn, 1941-1958
	- Harry Weyher, Jr., 1958-2002
	- J. Phillippe Rushton, 2002-12
	- Richard Lynn, 2012-??
 -->

<!--
## *Brown v Board* (1954) and *Mankind Quarterly*

:::: columns
::: {.column width="32%"}
![[Henry Garrett (1894-1973)](https://en.wikipedia.org/wiki/Henry_Garrett_(psychologist)). [Source](https://www.psychometricsociety.org/post/past-present-and-incoming-presidents), citing American Psychologist (1946), 1(9), 371](img/henry_e_garrett.png){width="100%"}
:::

::: {.column width="32%"}
![[Reginald Ruggles Gates (1882-1962)](https://en.wikipedia.org/wiki/Reginald_Ruggles_Gates), 1921. [Source](https://en.wikipedia.org/wiki/File:Reginald_Ruggles_Gates_Portrait.jpg)](img/Reginald_Ruggles_Gates_Portrait.jpg)
:::

::: {.column width="32%"}
![[George Robert Gayre of Gayre and Nigg (1907-1996)](https://en.wikipedia.org/wiki/Robert_Gayre). [Source](https://www.patrickcomerford.com/2020/07/robert-gair-racist-from-rathmines-who.html), original uncredited](img/Robert Gayre profile.jpg)
:::
::::

> Ultimately, anti-'racial' science played a defining role in the Supreme Court’s dismissal of the 'separate-but-equal' principle in education .... [@SchafferScientificRacismAgain2007 261; see also @JacksonScienceSegregationRace2005]

## *MQ* was founded as race science

> [[E]veryone agrees that races or ethnic groups exist and that they are in fact the raw material by means of which human evolution has taken place]{.red} ....

> [A scientific approach to racial questions, free from sentiment, propaganda or bias]{.blue}, may lead to a more reasonable attitude to the problems which agitate considerable sections of mankind. This need for a calmer approach to problems of race on the part of [civilised peoples as well as more primitive tribes]{.red} makes it desirable that a journal should exist in which the problems of racial origin and racial relationships can be quietly discussed, [without rancour or bigotry and with the primary aim of elucidating facts]{.blue}. (Gates, *MQ* 1:1 [1960], 11)

-->

# Mainstreaming race science {visibility="uncounted"}

## Pioneer-funded researchers

::::: columns
:::: column
::: {.smaller}
| | |
|:-----|:-----|
| [Thomas J. Bouchard, Jr.](https://en.wikipedia.org/wiki/Thomas_J._Bouchard_Jr.) | psychology |
| Brunetto Chiarelli | anthropology? |
| [Hans Eysenck](https://en.wikipedia.org/wiki/Hans_Eysenck) | psychology |
| [Robert Gordon](https://en.wikipedia.org/wiki/Robert_A._Gordon) | sociology |
| [Linda Gottfredson](https://en.wikipedia.org/wiki/Linda_Gottfredson) | psychology |
| [Garrett Hardin](https://en.wikipedia.org/wiki/Garrett_Hardin) | ecology |
| [Joseph M. Horn](https://en.wikipedia.org/wiki/Joseph_M._Horn) | psychology |
| [Lloyd Humphreys](https://en.wikipedia.org/wiki/Lloyd_Humphreys) | psychology |
| [Arthur Jensen](https://en.wikipedia.org/wiki/Arthur_Jensen) | psychology |
| [Michael Levin](https://en.wikipedia.org/wiki/Michael_Levin) | philosophy |
| [Richard Lynn](https://en.wikipedia.org/wiki/Richard_Lynn) | psychology |
| [R. Travis Osborne](https://en.wikipedia.org/wiki/R._Travis_Osborne) | psychology |
| [J. Phillippe Rushton](https://en.wikipedia.org/wiki/J._Phillippe_Rushton) | psychology |
| [Audrey M. Shuey](https://en.wikipedia.org/wiki/Audrey_M._Shuey) | psychology |
| [Philip A. Vernon](https://en.wikipedia.org/wiki/Philip_A._Vernon) | psychology |
| [Daniel Vining, Jr.](https://en.wikipedia.org/wiki/Daniel_Vining_Jr.) | demography |
:::
::::

:::: column
- @MillerPioneerFundBankrolling1994
- [List of Grantees from PF website](https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html)

\

- WoS author search results for 14 authors (Sept/Oct 2021)
- Identify journals that published more than 5/14
::::
:::::

## Pioneer-funded researchers

![Journals publishing 6 or more Pioneer-funded researchers, WOS author search results](img/wos_results.png)

## Text-mining corpus

|                                |       |
|:-------------------------------|------:|
| Behavior Genetics              |  2268 |
| Behavioral & Brain Sciences    |   898 |
| Intelligence                   |  1237 |
| **Mankind Quarterly**          |  1821 |
| Personality & Individual<br> Differences |  7274 |
| Psychological Reports          | 21398 |

- 34k journal articles
- Published 1960-2010

::: {.aside .smaller}
- 3 journals published by American Psychological Association not available
- In B&BS, focal papers + commentaries all bundled together
- PID is missing many relevant book reviews and essays, eg, [Rushton's review of Lynn's book on Pioneer](https://www.sciencedirect.com/science/article/pii/S0191886901001337)
:::

## Topic modeling

:::: columns
::: {.column width="60%"}
![Diachronic evolution of topic categories in *Philosophy of Science*, 1934-2015. @MalaterreWhatThisThing2019 fig. 2.](img/MalaterreWhatThisThing2019.jpeg){width="100%"}
:::

::: {.column width="40%"}
- Statistical method for clustering documents and words simultaneously

\

- [Can we identify a race science topic?]{.blue}
- [Does this topic start in MQ, spread to mainstream journals?]{.blue}
:::
::::

::: aside
- Custom R package [`tmfast`](https://github.com/dhicks/tmfast)
- 18 different models (3 vocabulary sizes $\times$ $k$ = 5, 10, 20, ..., 50)
:::

## UMAP interactive

```{r}
#| out-width: 1.1*1440px
#| out-height: 1080px
#| fig-cap: "UMAP visualization of Hellinger similarity, medium vocabulary, k = 30"

umap_vis = read_rds(here(vis_dir, '07_md_30.Rds'))
umap_vis
```

## 3 focal topics {.smaller}

```{r}
beta_gt = read_rds(here(vis_dir, '??_beta_table.Rds'))
beta_gt
```

## Focal topics, by journal, over time

![](img/08_presentation.png)

# Conclusions {visibility="hidden"}
## Conclusions {.smallish}

- Most Pioneer-funded researchers did not publish heavily in *MQ*
- *MQ* was the primary venue for one of two race science topics
- The other race science topic did not originate in *MQ*

. . .

\

- *MQ* was the primary venue for one form of race science
	- "Anthropology-flavored"? 
- While mainstream journals were the primary venues for a distinct form of race science
	- *Psychological Reports*, later *Intelligence* and *PID*
	- "Psychology-flavored"? 
- "The race science is coming from inside the house!"

# Thanks! {visibility="uncounted"}

<https://mainstream-race-science.netlify.app>

# Bonus content {visibility="uncounted"}

## Different topics or different communities?  {visibility="uncounted"}

![Authors by topic](img/_authors.png)

## Gould's critique of factor analysis {visibility="uncounted"}

> As a tool for simplification, [PCA/factor analysis] has proved its great value in many disciplines. But many factorists have gone beyond simplification, and tried to define factors as causal entities. This error of reification has plagued the technique since its inception .... [F]actors, by themselves, are neither things nor causes; they are mathematical abstractions. [@GouldMismeasureManRevised1996 284-5]

- `tmfast` uses methods from PCA and factor analysis [@RoheVintageFactorAnalysis2020]
	- (partial) PCA
	- varimax rotation

- Need to avoid hasty reification of topics

## Topic modeling as phenomena construction {visibility="uncounted"}

Topics as phenomena [@BogenSavingPhenomena1988; @WoodwardDataPhenomenaRestatement2009]

- simplified patterns extracted from word occurrence data
- may or may not be stable:\
  will we see the same topics in a different corpus? [@MalaterreEarlyDaysContemporary2022]
- constructed with minimal role for substantive theory
- **explananda, not explanans**

These words tend to be used together; **but why?**\
These test items tend to be answered in the same way; **but why?**

## References {visibility="uncounted"}