### 2023-06-09 paper/supplement_paper.tex
building paper infrastructure

% Use only LaTeX2e, calling the article.cls class and 12-point type.
\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

% Science template required packages
\usepackage{scicite}

\usepackage{times}

% add-ons for rticles
% based on arxiv.sty
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

% the basics
\usepackage{amsmath}
\usepackage{calc}
\usepackage{tabularx}

% for figure adjustment
\usepackage{placeins}
\usepackage{flafter}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.
\usepackage{setspace}

\newenvironment{sciabstract}{%
\begin{quote} \singlespacing}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{\bf Mainstreaming Race Science}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author{
Daniel J. Hicks,\textsuperscript{*}
and Emilio Lobato
\\\textsuperscript{*}To whom correspondence should be addressed; E-mail:
\href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}
\maketitle
\section*{Supplementary Text}
\renewcommand{\thesection}{S\arabic{section}}
\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

Appendix figures and tables will also need to be moved. A basic example
is provided here so this feature is also tested.

First we have a figure.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/thrice-1.pdf}
\caption{test thrice}
\end{figure}

And next we have a table

\begin{table}[ht]
\begin{center}
\begin{tabular}{ll}
thrice & 1 \\
thrice & 2\\
\end{tabular}
\end{center}
\end{table}

\end{document}