### 2023-07-19 paper/dataset.qmd
mostly filling in supplement refs

```{mermaid}
%%| fig-cap: Dataset construction and analysis overview
%%| fig-label: fig-corpus
graph TD

subgraph Mankind Quarterly Corpus
MQ("Mankind Quarterly<br>
    All articles, 1960-2004")
unz(Unz Review) --> MQ
pq(ProQuest) --> MQ
end

subgraph Pioneer Fund Corpus
au("Pioneer-funded authors<br>16") --> wos("Web of Science<br>
                                         author records")
wos --> journals
journals("Journals publishing<br> 
          6+ PF authors")
journals --> apa("APA journals<br>
                  not available")
journals --> generalist("Science & Nature<br>
                         excluded")
journals --> PF("Pioneer Fund Corpus<br>
                 5 journals<br>
                 All articles, 1960-2010")
end

subgraph Publishable Data and Code

MQ & PF --> corpus["Combined corpus<br>
                     (34k documents)"]
corpus --> pos("Noun phrase extraction")
pos --> vocab("Vocabulary selection<br>
                   (large, medium, small)")
vocab --> tm("Topic modeling<br>
               (k = 5, 10, 20, 30, 40, 50)")
tm --> space("Discursive space<br>analysis")
end
```