### 2024-05-03 paper/test.qmd
prep to merge back into master

This section uses simulated data to provide a proof-of-concept for our use of topic models to detect the "mainstreaming" of a marginalized topic.

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
theme_set(theme_bw())
library(tmfast)

library(glue)
library(tictoc)
library(memoise)

set.seed(2024-03-18)

## Simulation parameters
k = 20            ## total number of topics
N = 100           ## papers per journal
vocab = 1000      ## vocabulary size
paper_size = 500  ## length of each paper

paper_size
```